---
title: "LLIN_Analses"
author: "Steven Gan"
date: "`r Sys.Date()`"
format: pdf
toc: true
---

\pagebreak

# Load the Data

```{r}
library(readr)
ig2.df <- read_csv("Burkina Faso_bio_ig2_master_m36_anonym.csv")
pn3.df <- read_csv("Burkina Faso_bio_pn3_master_m36_anonym.csv")
interceptor.df <- read_csv("Burkina Faso_bio_interceptor_master_m36_anonym.csv")
```

# IG2 Data Preliminary Analyses

## Strucutre of IG2 Data

```{r}
colnames(ig2.df)
```

```{r}
table(ig2.df$NetType, ig2.df$Mosquitospeciesstrain)
```

```{r}
# An.kisumu is the same as An.gambiae Kisumu in "Mosquitospeciesstrain". 
# Merge them.
ig2.df$Mosquitospeciesstrain[
  ig2.df$Mosquitospeciesstrain == "An. gambiae Kisumu"] <- "An. Kisumu"
ig2.df$Mosquitospeciesstrain[
  ig2.df$Mosquitospeciesstrain == "An.kisumu"] <- "An. Kisumu"
# Untreated and Untreated net are the same as in "NetType".
ig2.df$NetType[ig2.df$NetType == "Untreated"] <- "Untreated Net"
ig2.df$NetType[ig2.df$NetType == "Untreated net"] <- "Untreated Net"
# One "NetType" is . and another is New. Remove it.
ig2.df <- ig2.df[ig2.df$NetType != ".", ]
ig2.df <- ig2.df[ig2.df$NetType != "New", ]

# Filter > 100 vales in "mort24h" and "mort72"
ig2.df <- ig2.df[ig2.df$mort24h <= 100, ]
ig2.df <- ig2.df[ig2.df$mort72 <= 100, ]

table(ig2.df$NetType, ig2.df$Mosquitospeciesstrain)
```


## Plot mortality at 24h and 72h v.s. round by type and mosquito species strain

```{r, warning=FALSE, message=FALSE, fig.width=8, fig.height=4}

# Plot "mort24h" & "mort72" v.s. "round" by "NetType" & "Mosquitospeciesstrain"

library(ggplot2)
library(gridExtra)

m1 <- ggplot(ig2.df, aes(x = round, y = mort24h)) +
  geom_point() + geom_smooth(method = "lm") +
  facet_grid(NetType ~ Mosquitospeciesstrain) +
  ylim(0, 100) +
  ylab("Mortality at 24h") + xlab("Round") +
  theme_bw()
m2 <- ggplot(ig2.df, aes(x = round, y = mort72)) +
  geom_point() + geom_smooth(method = "lm") +
  facet_grid(NetType ~ Mosquitospeciesstrain) +
  ylim(0, 100) +
  ylab("Mortality at 72h") + xlab("Round") +
  theme_bw()

grid.arrange(m1, m2, ncol = 2, 
             top = "Mortality vs Round of Ig2 Data")

ggsave("1_IG2_Mortality_vs_Round.png", width = 8, height = 4)
```

## Plot mortality at 24h and 72h v.s. drug conc. by round and mosquito species strain

```{r, warning=FALSE, message=FALSE, fig.width=8, fig.height=5}

# Plot "mort24h" v.s. "alpha_g_kg" & "chlor_g_kg" by "round" & "Mosquitospeciesstrain".

n1 <- ggplot(ig2.df, aes(x = alpha_g_kg, y = mort24h)) +
  geom_point() + geom_smooth() +
  facet_grid(round ~ Mosquitospeciesstrain) +
  ylim(0, 100) +
  ylab("Mortality at 24h") + xlab("Alpha concentration (g/kg)") +
  theme_bw()
n2 <- ggplot(ig2.df, aes(x = chlor_g_kg, y = mort24h)) +
  geom_point() + geom_smooth() +
  facet_grid(round ~ Mosquitospeciesstrain) +
  ylim(0, 100) +
  ylab("Mortality at 24h") + xlab("Chlor concentration (g/kg)") +
  theme_bw()

grid.arrange(n1, n2, ncol = 2, 
             top = "Mortality vs Drug Concentration of IG2 Data")

ggsave("2_IG2_Mortality_24h_vs_Drug_Concentration.png", width = 8, height = 5)
```

```{r, warning=FALSE, message=FALSE, fig.width=8, fig.height=3}

# Plot "mort24h" v.s. "alpha_g_kg" & "chlor_g_kg" by "Mosquitospeciesstrain".

n1.2 <- ggplot(ig2.df, aes(x = alpha_g_kg, y = mort24h)) +
  geom_point() + geom_smooth() +
  facet_wrap(~ Mosquitospeciesstrain) +
  ylim(0, 100) +
  ylab("Mortality at 24h") + xlab("Alpha concentration (g/kg)") +
  theme_bw()
n2.2 <- ggplot(ig2.df, aes(x = chlor_g_kg, y = mort24h)) +
  geom_point() + geom_smooth() +
  facet_wrap(~ Mosquitospeciesstrain) +
  ylim(0, 100) +
  ylab("Mortality at 24h") + xlab("Chlor concentration (g/kg)") +
  theme_bw()

grid.arrange(n1.2, n2.2, ncol = 2, 
             top = "Mortality vs Drug Concentration of IG2 Data")

ggsave("3_IG2_Mortality_vs_Drug_Concentration.png", width = 8, height = 3)
```

# PN3 Data Preliminary Analyses

```{r}
colnames(pn3.df)
```

```{r}
table(pn3.df$NetType, pn3.df$Mosquitospeciesstrain)
```

```{r}
# Change An. gambiae Kisumu to An. kisumu in "Mosquitospeciesstrain".
pn3.df$Mosquitospeciesstrain[
  pn3.df$Mosquitospeciesstrain == "An. gambiae Kisumu"] <- "An. Kisumu"
# Contrôle is the same as Untreated Net_Control  in "NetType".
# Merge them and name as "Untreated Net".
pn3.df$NetType[pn3.df$NetType == "Contrôle"] <- "Untreated Net"
pn3.df$NetType[pn3.df$NetType == "Untreated Net_Control"] <- "Untreated Net"

# Filter > 100 vales in "mort" and "kd60"
pn3.df <- pn3.df[pn3.df$mort <= 100, ]
pn3.df <- pn3.df[pn3.df$kd60 <= 100, ]

# Remove NA in "NetsType" and "Mosquitospeciesstrain"
pn3.df <- pn3.df[!is.na(pn3.df$NetType), ]
pn3.df <- pn3.df[!is.na(pn3.df$Mosquitospeciesstrain), ]

table(pn3.df$NetType, pn3.df$Mosquitospeciesstrain)
```

## Plot knock out rate at 60 min and moratality v.s. round by type and mosquito species strain

```{r, warning=FALSE, message=FALSE, fig.width=8, fig.height=4}
# Plot "kd60" & "mort" v.s. "round" by "NetType" & "Mosquitospeciesstrain"

o1 <- ggplot(pn3.df, aes(x = round, y = kd60)) +
  geom_point() + geom_smooth(method = "lm") +
  facet_grid(NetType ~ Mosquitospeciesstrain) +
  ylim(0, 100) +
  ylab("Knock Down Rate at 60 min") + xlab("Round") +
  theme_bw()
o2 <- ggplot(pn3.df, aes(x = round, y = mort)) +
  geom_point() + geom_smooth(method = "lm") +
  facet_grid(NetType ~ Mosquitospeciesstrain) +
  ylim(0, 100) +
  ylab("Mortality") + xlab("Round") +
  theme_bw()

grid.arrange(o1, o2, ncol = 2, 
             top = "Knock Down Rate and Mortality vs Round of PN3 Data")

ggsave("4_PN3_Knock_Down_Rate_and_Mortality_vs_Round.png", 
       width = 8, height = 4)
```


## Plot knock out rate at 60 min and mortality v.s. drug conc. by round and mosquito species strain

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=5}
# Plot "kd60" v.s. "delta_side_g_kg" & "delta_roof_g_kg" & "pbo_roof_g_kg" 
# by "round" & "Mosquitospeciesstrain".

p1 <- ggplot(pn3.df, aes(x = delta_side_g_kg, y = kd60)) +
  geom_point() + geom_smooth() +
  facet_grid(round ~ Mosquitospeciesstrain) +
  ylim(0, 100) +
  ylab("Knock Down Rate at 60 min") + xlab("Delta Side concentration (g/kg)") +
  theme_bw()
p2 <- ggplot(pn3.df, aes(x = delta_roof_g_kg, y = kd60)) +
  geom_point() + geom_smooth() +
  facet_grid(round ~ Mosquitospeciesstrain) +
  ylim(0, 100) +
  ylab("Knock Down Rate at 60 min") + xlab("Delta Roof concentration (g/kg)") +
  theme_bw()
p3 <- ggplot(pn3.df, aes(x = pbo_roof_g_kg, y = kd60)) +
  geom_point() + geom_smooth() +
  facet_grid(round ~ Mosquitospeciesstrain) +
  ylim(0, 100) +
  ylab("Knock Down Rate at 60 min") + xlab("PBO Roof concentration (g/kg)") +
  theme_bw()

grid.arrange(p1, p2, p3, ncol = 3, 
             top = "Knock Down Rate vs Drug Concentration of PN3 Data")

ggsave("5_PN3_Knock_Down_Rate_vs_Drug_Concentration.png", 
       width = 12, height = 5)
```

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=5}

# Plot "mort" v.s. "delta_side_g_kg" & "delta_roof_g_kg" & "pbo_roof_g_kg"
# by "round" & "Mosquitospeciesstrain".

q1 <- ggplot(pn3.df, aes(x = delta_side_g_kg, y = mort)) +
  geom_point() + geom_smooth() +
  facet_grid(round ~ Mosquitospeciesstrain) +
  ylim(0, 100) +
  ylab("Mortality") + xlab("Delta Side concentration (g/kg)") +
  theme_bw()
q2 <- ggplot(pn3.df, aes(x = delta_roof_g_kg, y = mort)) +
  geom_point() + geom_smooth() +
  facet_grid(round ~ Mosquitospeciesstrain) +
  ylim(0, 100) +
  ylab("Mortality") + xlab("Delta Roof concentration (g/kg)") +
  theme_bw()
q3 <- ggplot(pn3.df, aes(x = pbo_roof_g_kg, y = mort)) +
  geom_point() + geom_smooth() +
  facet_grid(round ~ Mosquitospeciesstrain) +
  ylim(0, 100) +
  ylab("Mortality") + xlab("PBO Roof concentration (g/kg)") +
  theme_bw()

grid.arrange(q1, q2, q3, ncol = 3, 
             top = "Mortality vs Drug Concentration of PN3 Data")

ggsave("6_PN3_Mortality_vs_Drug_Concentration.png", width = 12, height = 5)
```

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=3}

# Plot "mort" v.s. "delta_side_g_kg" & "delta_roof_g_kg" & "pbo_roof_g_kg"
# by "Mosquitospeciesstrain".

q1.2 <- ggplot(pn3.df, aes(x = delta_side_g_kg, y = mort)) +
  geom_point() + geom_smooth() +
  facet_wrap(~ Mosquitospeciesstrain) +
  ylim(0, 100) +
  ylab("Mortality") + xlab("Delta Side concentration (g/kg)") +
  theme_bw()
q2.2 <- ggplot(pn3.df, aes(x = delta_roof_g_kg, y = mort)) +
  geom_point() + geom_smooth() +
  facet_wrap(~ Mosquitospeciesstrain) +
  ylim(0, 100) +
  ylab("Mortality") + xlab("Delta Roof concentration (g/kg)") +
  theme_bw()
q3.2 <- ggplot(pn3.df, aes(x = pbo_roof_g_kg, y = mort)) +
  geom_point() + geom_smooth() +
  facet_wrap(~ Mosquitospeciesstrain) +
  ylim(0, 100) +
  ylab("Mortality") + xlab("PBO Roof concentration (g/kg)") +
  theme_bw()

grid.arrange(q1.2, q2.2, q3.2, ncol = 3,
             top = "Mortality vs Drug Concentration of PN3 Data")

ggsave("7_PN3_Mortality_vs_Drug_Concentration.png", width = 12, height = 3)
```

# Comparison of IG2 and PN3 Data

## Data structure

```{r}
library(dplyr)

# Select An. coluzzii VKPER from "Mosquitospeciesstrain" in ig2.df.
# Select Interceptor G2 from "NetType" in ig2.df.
# Extract "NetType", "round", and "mort24h" from ig2.df
# Name 3 columns as "NetType", "round", and "mort" respectively.
ig2.mort <- ig2.df %>%
  filter(Mosquitospeciesstrain == "An. coluzzii VKPER") %>%
  filter(NetType == "Interceptor G2") %>%
  select(NetType, round, mort24h) %>%
  rename(mort = mort24h)

# Select An. coluzzii VKPER from "Mosquitospeciesstrain" in pn3.df.
# Select PermaNet 3.0 from "NetType" in pn3.df.
# Extract "NetType", "round", and "mort" from pn3.df
pn3.mort <- pn3.df %>%
  filter(Mosquitospeciesstrain == "An. coluzzii VKPER") %>%
  filter(NetType == "PermaNet 3.0") %>%
  select(NetType, round, mort) 

# Combine ig2.mort and pn3.mort
ig2.pn3.mort <- rbind(ig2.mort, pn3.mort)
```

## Dot plot Mortality Comparison of IG2 and PN3 Data

```{r, warning=FALSE, message=FALSE}
# Plot "mort" v.s. "round" by "NetType" of ig2.pn3.mort.

ggplot(ig2.pn3.mort, aes(x = round, y = mort, color = NetType,)) +
  geom_point() + geom_smooth(method = "lm") +
  facet_wrap(~ NetType) +
  ylim(0, 100) +
  ylab("Mortality") + xlab("Round") +
  theme_bw() +
  ggtitle("Mortality Comparison of IG2 and PN3 Data")

ggsave("8_Mortality_Comparison_of_IG2_and_PN3_Data.png", width = 6, height = 3)
```

## Statistical inference

```{r}
# Perform t-test to compare mortality between IG2 and PN3 for each round.
# t.test ~ 2 independent samples
cat("t-test: IG2 vs PN3\n")
cat("p-value for each round:\n")
t.test(mort ~ NetType, ig2.pn3.mort[ig2.pn3.mort$round == 0, ])$p.value
t.test(mort ~ NetType, ig2.pn3.mort[ig2.pn3.mort$round == 1, ])$p.value
t.test(mort ~ NetType, ig2.pn3.mort[ig2.pn3.mort$round == 2, ])$p.value
t.test(mort ~ NetType, ig2.pn3.mort[ig2.pn3.mort$round == 3, ])$p.value

cat("\n")

# Perform U-test to compare mortality between IG2 and PN3 for each round.
# wilcox.test ~ 2 independent samples
cat("U-test: IG2 vs PN3\n")
cat("p-value for each round:\n")
wilcox.test(mort ~ NetType, ig2.pn3.mort[ig2.pn3.mort$round == 0, ])$p.value
wilcox.test(mort ~ NetType, ig2.pn3.mort[ig2.pn3.mort$round == 1, ])$p.value
wilcox.test(mort ~ NetType, ig2.pn3.mort[ig2.pn3.mort$round == 2, ])$p.value
wilcox.test(mort ~ NetType, ig2.pn3.mort[ig2.pn3.mort$round == 3, ])$p.value
```

## Violin plot Mortality Comparison of IG2 and PN3 Data

```{r, warning=FALSE, message=FALSE}

# Plot "mort" v.s. "round" of ig2.pn3.mort by "NetType".
library(ggpubr)

ggviolin(ig2.pn3.mort, x = "round", y = "mort", color = "NetType",
         palette = "jco", add = "mean_sd", error.plot = "crossbar",
         add.params = list(size = 0.5)) +
  geom_jitter(aes(color = NetType), position = position_jitterdodge(0.2), 
              size = 0.5) +
  theme_bw() + ylim(0, 100) +
  labs(x = "Round", y = "Mortality",
       title = "Mortality Comparison of IG2 and PN3 Data",
       subtitle = "An. coluzzii VKPER") +
  # U test of 2 independent samples for each round
  stat_compare_means(method = "wilcox.test", label = "p.signif",
                     aes(group = NetType), label.y = 99)

ggsave("9_Mortality_Comparison_of_IG2_and_PN3_Data_Violin.png", 
       width = 6, height = 4)
```

## Box plot Mortality Comparison of IG2 and PN3 Data

```{r, warning=FALSE, message=FALSE}
# Plot "mort" v.s. "round" of ig2.pn3.mort by "NetType".

ggboxplot(ig2.pn3.mort, x = "round", y = "mort", color = "NetType",
          palette = "jco") +
  geom_jitter(aes(color = NetType), position = position_jitterdodge(0.2),
              size = 1) +
  theme_bw() + ylim(0, 100) +
  labs(x = "Round", y = "Mortality",
       title = "Mortality Comparison of IG2 and PN3 Data",
       subtitle = "An. coluzzii VKPER") +
  # U test of 2 independent samples for each round
  stat_compare_means(method = "wilcox.test", label = "p.signif",
                     aes(group = NetType), label.y = 99)

ggsave("10_Mortality_Comparison_of_IG2_and_PN3_Data_Box.png", 
       width = 5, height = 3)
```



# Interceptor Data Preliminary Analyses

## Structure of Interceptor Data

```{r}
colnames(interceptor.df)
```

```{r}
table(interceptor.df$Mosquitospeciesstrain)
```

